---
title: "Assignment 1: Application"
author: "Lien Dao"
date: "2/27/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Indicators data
```{r}
library(datasets)
library(readr)
url <- "https://gattonweb.uky.edu/sheather/book/docs/datasets/indicators.txt"
indicators <- read.table(url, header = TRUE)
print(indicators)
```
\break

## Question 6
**Simple linear regression model**
```{r}
indicators_lm <- lm(PriceChange ~ LoanPaymentsOverdue, data = indicators)
summary(indicators_lm)
```
\hfill
a) 95% CI
```{r}
confint_95 <- confint(indicators_lm)[2, ]
confint_95
```
\hfill
b) 95% CI for E[Y|X=4]
```{r}
predict(indicators_lm, data.frame(LoanPaymentsOverdue = 4), interval = "confidence")
```
\break
## Question 7
**Function to calculate basic statistics**
```{r} 
RSS <- function(x,y){
  mean_x <- mean(x)
  mean_y <- mean(y)
  
  Sxy = sum((x - mean(x)) * (y - mean(y)))
  Sxx = sum((x - mean(x))^2)
  Syy = sum((y - mean(y))^2)
    
  beta_1_hat = Sxy / Sxx
  beta_0_hat = mean(y) - beta_1_hat * mean(x)
  result <- data.frame(mean_x, mean_y,
                       Sxx, Syy, Sxy, 
                       beta_0_hat, beta_1_hat)
}
```
\hfill
**Results**
```{r}
rss_all <- RSS(indicators$LoanPaymentsOverdue, indicators$PriceChange)
rss_all
```
\break
## Question 8
a) Calculate predicted y
```{r}
indicators$pred_y <- rss_all[1,6] + rss_all[1,7]*indicators$LoanPaymentsOverdue
```
\hfill
**Function to get ANOVA table's elements**
```{r}
ANOVA <- function(n,x,y,y_pred){
  df_reg <- 1
  df_res <- n-2
  df_total <- n-1

  SS_reg = sum((y_pred - mean(y))^2)
  SSE = sum((y - y_pred)^2)
  SST = sum((y - mean(y))^2)
  
  MSR = SS_reg/df_reg
  MSE = SSE/df_res
  
  F_stat = MSR/MSE
  
  p_value <- pf(F_stat, df_reg, df_res, lower.tail = FALSE)
  result <- data.frame(df_reg, df_res, df_total,
                       SS_reg, SSE, SST,
                       MSR, MSE, NA,
                       F_stat, NA, NA,
                       p_value, NA, NA)
}
```
\hfill
**ANOVA table**
```{r}
num_value <- 18
anova_all <- ANOVA(num_value,indicators$LoanPaymentsOverdue,indicators$PriceChange,indicators$pred_y)
anova_table <- matrix(anova_all,ncol = 5)
dimnames(anova_table) <- list("  Group" = c("Regression", "Residual", "Total"),
                              "ANOVA" = c("Df", "SS", "MS", "F stat", "P value"))
anova_table
```
\hfill
b) Unbiased estimator of Var(Y|X) is S^2 (or MSE)
```{r}
unbiased_s <- anova_all[1,8]
unbiased_s
```
\break
## Question 9
a) Standard error of b1 hat 
```{r}
se_b1_hat <- sqrt(unbiased_s/rss_all[1,3])
se_b1_hat
```
\hfill
b) Ho: b1=0; Ha: b1 >< 0
```{r}
t_stat <- abs(rss_all[1,7]/se_b1_hat)
t_stat
t_critical <- qt(p=0.05/2, df=(num_value-2), lower.tail=FALSE)
t_critical
as.logical(t_stat>t_critical)
```
\hfill
*Conclusion:* We reject the null hypothesis and conclude that there a linear association between X and Y at 0.05 significant level

\break

## Question 10
a) 95% CI for b1
```{r}
lwb <- rss_all[1,7] - t_critical*se_b1_hat
upb <- rss_all[1,7] + t_critical*se_b1_hat
confint_b1 <- c(lwb, upb)
confint_b1
```
\hfill
*Conclusion:* We reject the null hypothesis because 0 does not belong to the confidence interval (we have the same conclusion as before)

\hfill

b) 95% CIs for population regression line and prediction intervals of Y

**Function to obtain 95% CIs for population regression line**
```{r}
confi_reg <- function(n,x_real,x,y){
  t_critical <- qt(p=0.05/2, df=(num_value-2), lower.tail=FALSE)
  y_pred <- rss_all[1,6] + rss_all[1,7]*x_real
  lwb <- y_pred - t_critical*sqrt(unbiased_s*((1/n)+(x_real-mean(x))^2/rss_all[1,3]))
  upb <- y_pred + t_critical*sqrt(unbiased_s*((1/n)+(x_real-mean(x))^2/rss_all[1,3]))
  result <- c(lwb,upb)
}
```
\hfill
x = 2.1
```{r}
x1 <- confi_reg(num_value,2.1,indicators$LoanPaymentsOverdue,indicators$PriceChange)
x1
```
\hfill
x = 3.3
```{r}
x2 <- confi_reg(num_value,3.3,indicators$LoanPaymentsOverdue,indicators$PriceChange)
x2
```
\hfill
x = 4.0
```{r}
x3 <- confi_reg(num_value,4.0,indicators$LoanPaymentsOverdue,indicators$PriceChange)
x3
```
\hfill
x = 4.4
```{r}
x4 <- confi_reg(num_value,4.4,indicators$LoanPaymentsOverdue,indicators$PriceChange)
x4
```
\hfill
**Function to obtain prediction intervals of Y**
```{r}
confi_Y <- function(n,x_real,x,y){
  t_critical <- qt(p=0.05/2, df=(num_value-2), lower.tail=FALSE)
  y_pred <- rss_all[1,6] + rss_all[1,7]*x_real
  lwb <- y_pred - t_critical*sqrt(unbiased_s*(1+(1/n)+(x_real-mean(x))^2/rss_all[1,3]))
  upb <- y_pred + t_critical*sqrt(unbiased_s*(1+(1/n)+(x_real-mean(x))^2/rss_all[1,3]))
  result <- c(lwb,upb)
}
```
\hfill
x = 2.1
```{r}
y_x1 <- confi_Y(num_value,2.1,indicators$LoanPaymentsOverdue,indicators$PriceChange)
y_x1
```
\hfill
x = 3.3
```{r}
y_x2 <- confi_Y(num_value,3.3,indicators$LoanPaymentsOverdue,indicators$PriceChange)
y_x2
```
\hfill
x = 4.0
```{r}
y_x3 <- confi_Y(num_value,4.0,indicators$LoanPaymentsOverdue,indicators$PriceChange)
y_x3
```
\hfill
x = 4.4
```{r}
y_x4 <- confi_Y(num_value,4.4,indicators$LoanPaymentsOverdue,indicators$PriceChange)
y_x4
```
\hfill
## End

